---
title: "House Prices_Gabriella"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
source('src/ini.R')
source('src/load.R')
source('src/transform.R')
source('src/impute.R')

```

## EDA - Categorical Variables

```{r}
source('src/feature_eng.R')

p = ggplot(train) + theme_bw()

SP_mean = mean(train$SalePriceLog)

draw_boxplot = function(var){
  p + geom_boxplot(aes_string(x = var, y = 'SalePriceLog')) + ggtitle(var) +
    geom_hline(yintercept=SP_mean, linetype="dashed", color = "red") 
}
plotlist = lapply(train, function(x){ifelse(class(x) %in% c('factor'), 1, NA)}) %>% unlist()
plotlist = names(plotlist)[which(plotlist == 1)]
plotlist = lapply(plotlist, function(x){draw_boxplot(x)})
multiplot(cols = 5, plotlist = plotlist)


```

In order to observe a correlation between the categorical features and SalesPrice(Log), we will consider the Kendall correlation. This method correlates as.numeric(factor) and shows relationships between features.
? Was dummy varibles used?

```{r}
# Chnaging Factor variables to Numeric (in order to develop cor matrix)
MSSubClass = as.numeric(train$MSSubClass)
MSZoning = as.numeric(train$MSZoning)
Street = as.numeric(train$Street)
Alley = as.numeric(train$Alley)
LotShape = as.numeric(train$LotShape)
LandContour = as.numeric(train$LandContour)
Utilities = as.numeric(train$Utilities)
LotConfig = as.numeric(train$LotConfig)
LandSlope = as.numeric(train$LandSlope)
Neighborhood = as.numeric(train$Neighborhood)
Condition1 = as.numeric(train$Condition1)
Condition2 = as.numeric(train$Condition2)
BldgType = as.numeric(train$BldgType)
HouseStyle = as.numeric(train$HouseStyle)
RoofStyle = as.numeric(train$RoofStyle)
RoofMatl = as.numeric(train$RoofMatl)
Exterior1st = as.numeric(train$Exterior1st)
Exterior2nd = as.numeric(train$Exterior2nd)
MasVnrType = as.numeric(train$MasVnrType)
ExterQual = as.numeric(train$ExterQual)
ExterCond = as.numeric(train$ExterCond)
BsmtQual = as.numeric(train$BsmtQual)
BsmtCond = as.numeric(train$BsmtCond)
BsmtExposure = as.numeric(train$BsmtExposure)
BsmtFinType1 = as.numeric(train$BsmtFinType1)
BsmtFinType2 = as.numeric(train$BsmtFinType2)
Heating = as.numeric(train$Heating)
HeatingQC = as.numeric(train$HeatingQC)
CentralAir = as.numeric(train$CentralAir)
Electrical = as.numeric(train$Electrical)
KitchenQual = as.numeric(train$KitchenQual)
FireplaceQu = as.numeric(train$FireplaceQu)
GarageType = as.numeric(train$GarageType)
GarageFinish = as.numeric(train$GarageFinish)
GarageQual = as.numeric(train$GarageQual)
GarageCond = as.numeric(train$GarageCond)
PavedDrive = as.numeric(train$PavedDrive)
PoolQC = as.numeric(train$PoolQC)
Fence = as.numeric(train$Fence)
MiscFeature = as.numeric(train$MiscFeature)
SaleType = as.numeric(train$SaleType)
SaleCondition = as.numeric(train$SaleCondition)
Functional = as.numeric(train$Functional)
Foundation = as.numeric(train$Foundation)
```

# Property Variables
```{r}
m = cbind(SalesLog = train$SalePriceLog, Street, Alley, LotShape, LandContour, LandSlope, LotConfig)
cor(m, method="kendall", use="pairwise") 

```
No variables have a significant relationship to SalesPriceLog. It might be worth considering LotShape.
It seems that only LandContour & LandSlope have a strong negative correlation. 

# Dwelling/ Living Variables
```{r}
m1 = cbind(SalesLog = train$SalePriceLog, BldgType, HouseStyle, MSSubClass)
cor(m1, method="kendall", use="pairwise") 

```
No variables have a significant relationship to SalesPriceLog.
It seems that only BldgType & MSSubClass have a strong positive correlation. 

# Utility Variables
```{r}
m2 = cbind(SalesLog = train$SalePriceLog, Utilities, Electrical)
cor(m2, method="kendall", use="pairwise") 

```
Electrical seems to correlate to SalesLog but not significantly. 

# Overall Quality & Condition Variables
```{r}
m3 = cbind(SalesLog = train$SalePriceLog, OverallQual= train$OverallQual, OverallCond = train$OverallCond, Functional)
cor(m3, method="kendall", use="pairwise") 

```
OverallQual has shown the strongest positive correlation to SalesPriceLog. 
Interestingly, I would have thought OverallQual & OverallCond were strongly related, but aren't. 

# Material of House Variables
```{r}
m4 = cbind(SalesLog = train$SalePriceLog, OverallQual = train$OverallQual, RoofMatl, Exterior1st, Exterior2nd, MasVnrType, ExterQual, ExterCond, Foundation)
cor(m4, method="kendall", use="pairwise") 

```
ExterQual seems to have a negative correlation to SalesPriceLog. 
ExterQual & OverallQual have a strong correlaton/ relationship. It might be worth considering only one of these factors. 
Additionally, Exterior1st strongly related to Exterior2nd.

# Roof Variables
```{r}
m5 = cbind(SalesLog = train$SalePriceLog, RoofStyle, RoofMatl)
cor(m5, method="kendall", use="pairwise") 

```

# Basement Variables
```{r}
m6 = cbind(SalesLog = train$SalePriceLog, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2)
cor(m6, method="kendall", use="pairwise") 

```
BsmtQual seems to have a relationship to SalesPriceLog. 

# Heating Variables
```{r}
m7 = cbind(SalesLog = train$SalePriceLog, Heating, HeatingQC, CentralAir, FireplaceQu, Neighborhood)
cor(m7, method="kendall", use="pairwise") 

```


# Garage Variables
```{r}
m8 = cbind(SalesLog = train$SalePriceLog, GarageType, GarageFinish, GarageQual, GarageCond, PavedDrive)
cor(m8, method="kendall", use="pairwise") 

```
GarageFinish seems to have a relationship to SalesPriceLog.

# Misc Variables
```{r}
m9 = cbind(SalesLog = train$SalePriceLog, PoolQC, Fence, MiscFeature)
cor(m9, method="kendall", use="pairwise") 

```

# Nature of Sale Variables
```{r}
m10 = cbind(SalesLog = train$SalePriceLog, SaleType, SaleCondition)
cor(m10, method="kendall", use="pairwise") 

```
SaleCondition & SaleType do have a relationship.
I would have thought that this would have had a stronger relationship. 

# Remaining Variables
```{r}
m11 = cbind(SalesLog = train$SalePriceLog, Neighborhood, KitchenQual, MSZoning, Condition1, Condition2)
cor(m11, method="kendall", use="pairwise") 

```
